---
title: "Class09"
author: "Kellie Lemoine"
date: "May 1, 2019"
output: github_document
---

##Import data
```{r}
fna.data <- "data/WisconsinCancer.csv"
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

##Determine number of patients with malignant tumors
```{r}
table(wisc.df$diagnosis)
```

##How many variables/features in the data are suffixed with _mean?
```{r}
length(grep("_mean", colnames(wisc.df)))
```

Select cols 3:32 and store it as a matrix
```{r}
wisc.data <- as.matrix(wisc.df[ , 3:32])
```

```{r}
row.names(wisc.data) <- wisc.df$id
```

```{r}
diagnosis <- wisc.df$diagnosis
```


Section 2: Principal Component Analysis (PCA)


##Check column means and standard deviations
```{r}
round(apply(wisc.data, 2, mean), 1)

round(apply(wisc.data, 2, sd), 1)
```

```{r}
wisc.pr <- prcomp(wisc.data, scale. = TRUE)

summary(wisc.pr)
```

##Generate a biplot of the data

```{r}
biplot(wisc.pr)
```

##Generate a clearer plot of the data
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis, xlab = "PC1", ylab = "PC2")
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, xlab = "PC1", ylab = "PC3")
```

##Calculate variance
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)

pve <- pr.var/0.4427
```

## Plot variance explained for each principal component
```{r}
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


##Communicating PCA results
```{r}
round(wisc.pr$rotation["radius_mean",1], 2)
```

```{r}
round(wisc.pr$rotation["smoothness_se",1], 2)
```

```{r}
sort(abs((wisc.pr$rotation[,1])))
```


Section 3: Hierarchial Clustering


```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)

abline(h = 19, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, h=19, k=4)
table(wisc.hclust.clusters)
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```


Section 5:Combining Methods

##I am going to start with the PCs that capture 90% of the original variance in the dataset (i.e. PC1 to PC7)

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method="ward.D2")

plot(wisc.pr.hclust)
```


```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```




























