---
title: 'Class 8: Machine learning'
author: "Kellie Lemoine"
date: "April 26, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Kmeans clustering

Let's try out the **kmeans()** function in R with some makeup data

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

Use the Kmeans() function setting k to 2 and nstart=20

```{r}
km <- kmeans(x, centers = 2, 20)
km
```


Inspect/print the results

> How many points are in each cluster? 

```{r}
km$size
table(km$cluster)
```


> Q. What ‘component’ of your result object details
 - cluster size? nstart
 - cluster assignment/membership?
 - cluster center?

```{r}
km$centers
```


Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points

```{r}
plot(x, col = km$cluster)
points(km$centers, col = "blue", pch = 18, cex = 3)
```

## Hierarchial Clustering

Here we don't have to spell out K the number of clusters before hand but we do have to give it a distance matrix

```{r}
d <- dist(x)
hc <- hclust(dist(x))
hc
```

Lets plot the results

```{r}
plot(hc)
abline(h=6, col="red")
cutree(hc, h= 6)
```

```{r}
gp2 <- cutree(hc, k= 2)
gp3 <- cutree(hc, k=2)
```


```{r}
table(gp2)
table(gp3)
```

```{r}
table(gp2, gp3)
```

Try a more real-life example to see how our clustering works

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

> Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters

```{r}
#hc <- hclust(dist(x))
#plot(hc)
#abline(1.8)
#cutree(hc, h = 2)
```

```{r}
plot(x, col = gp3)
```


> Q. How does this compare to your known 'col' groups?



Principal Component Analysis (PCA)

We will use the base R **prcomp()** function for PCA today...

Let's fet some RNASeq data to play with

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)

head(mydata) 
```

There are `r `nrow(mydata)` genes in this dataset

```{r}
pca <- prcomp( t(mydata), scale=TRUE)
summary(pca) 
```

```{r}
attributes(pca)
```

Let's make our first PCA plot

```{r}
plot(pca$x[,1], pca$x[,2])
```

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab = paste("PC1 (",pca.var.per[1],"%)", sep=""))
```

```{r}
xlab <- paste("PC1 (",pca.var.per[1],"%)", sep="")
ylab <- paste("PC2 (",pca.var.per[2],"%)", sep="")

xlab
ylab
```

```{r}
mycols <- c( rep("red", 5), rep("blue", 5))
```


```{r}
plot(pca$x[,1], pca$x[,2], xlab=xlab, ylab=ylab, col=mycols)
text(pca$x[,1], pca$x[,2], colnames(mydata))
```


## Hands on activity

```{r}

x <- read.csv("UK_foods.csv", row.names = 1)

nrow(x)
ncol(x)

```

```{r}
## call view function

#View(x)
```

```{r}
## Check dimensions

dim(x)
```

```{r}
## Generate a barplot of the data

barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
mycols <- c("orange", "red", "blue", "green")
plot(pca$x[,1], pca$x[,2])
text(pca$x[,1], pca$x[,2], colnames(x), col = mycols)
abline(h=0, col="gray", lty=2)
```
















