---
title: "Class12"
author: "Kellie Lemoine"
date: "May 10, 2019"
output: github_document
---

## Load bio3d and read 1HSG files

```{r}
library("bio3d")

file <- get.pdb("1hsg")

pdb <- read.pdb(file)
```

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Trim files to protein and ligand only files and convert to PDB files

```{r}
##Select protein only

prot <- atom.select(pdb, "protein", value = TRUE)

prot

write.pdb(prot, file = "1hsg_protein.pdb")
```

```{r}

##Select ligand only
lig <- atom.select(pdb, "ligand", value = TRUE)

lig

write.pdb(lig, file = "1hsg_ligand.pdb")
```

## Add hydrogens and charges in ADT

We opened out protein only PDB file in AutoDoc tools and added hydrogens and atom-types needed for docking calculations

## Run docking

We will use AutoDoc Vina here at the UNIX command line!

##Process results back into R

```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
res
write.pdb(res, file="results.pdb")
```


 Compare our docking poses to the MERK drug structure
 
```{r}
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```
 
 ## Normal Mode Analysis example
 
```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```

```{r}
plot(modes)
```























