---
title: "Class 14:Transcriptomics and the analysis of RNA-Seq data"
author: "Kellie Lemoine"
date: "May 17, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


```{r}
#install.packages("BiocManager")
```

Import our countdata and coldata into R

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

Q1. How many genes are we dealing with here?

```{r}
nrow(counts)
```

Let's check if the cols of 'counts' is the same as the id column in 'metadata'

```{r}
colnames(counts)
```

```{r}
metadata$id
```

```{r}
all(colnames(counts)[-1] == metadata$id)
```

Let's access just the control counts data so we can calculate an average counts value for eacch gene in these control samples.

```{r}
control <- metadata[metadata[,"dex"]=="control",]
control
```

Now we want to get the cols in counts with these 'control$id' colnames

```{r}
control.mean <- rowSums( counts[ ,control$id] )/length(control$id)
names(control.mean) <- counts$ensgene
head(control.mean)
```

Now do the same for the 'treated' samples...

```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated
```

```{r}
treated.mean <- rowSums( counts[ ,treated$id] )/length(treated$id)
names(treated.mean) <- counts$ensgene
```


Let's put these together into a new data.frame called 'meancounts'

```{r}
meancounts <- data.frame(control.mean, treated.mean)
head(meancounts)
```

```{r}
plot(meancounts)
```

Plot on a log scale for both axis

```{r}
plot(meancounts, log="xy")
```

Calculate fold change and use log2 values

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

Remove the NaN and -inf

```{r}
#Identify the rows (i.e. genes) with xero values
to.rm <- unique(which(meancounts[,1:2]==0, arr.ind = TRUE)[,1])

#Remove these rows as we can't really say much about them
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.

Q8. Using the up.ind vector above can you determine how many up regulated genes we have at the greater than 2 fc level?

```{r}
up.ind <- mycounts$log2fc > 2
sum(up.ind)
```

```{r}
down.ind <- mycounts$log2fc < (-2)
sum(down.ind)
```

Adding annotation data

Read annotation data from ENSEMBLE...

```{r}
anno <- read.csv("annotables_grch38.csv")
```

We will use the **merge()** function to add annotation data to our 'mycounts' results

```{r}
mycounts.anno <- merge(x=mycounts, y=anno, by.x="row.names", by.y="ensgene")
```

We can use bioconductor to do this too

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```

```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")

mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")

head(mycounts)
```

```{r}
View( mycounts[up.ind,] )
```


Section 5: DESeq2 analysis

Let's do this the right way/ DESeq2 is an R package for analyzing count-based NGS data like RNA-seq. It is available from bioconductor

```{r}
library("DESeq2")
```

```{r}
citation("DESeq2")
```

Build our deseq dataset object need for running deseq analysis

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
#dds
```

```{r}
dds <- DESeq(dds)
```

Get our results

```{r}
res <- results(dds)
#as.data.frame(res)
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha = 0.05)
summary(res05)
```

```{r}
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange,  -log(res$padj), col=mycols, 
 ylab="-Log(P-value)", xlab="Log2(FoldChange)")

abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.05), col="darkgray", lty=2)

```

Another way to plot this is with the Enhanced

But first lets putproper gene names on our results (i.e. gene 'symbols')

```{r}
x <- as.data.frame(res)
x$symbol <- mapIds(org.Hs.eg.db, 
                   keys=row.names(x),
                   keytype="ENSEMBL",
                   column="SYMBOL",
                   multiVals="first")
```











